# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
(1/2) We load the data and set the time zone, but make no transformations:
```{r echo=TRUE}
## Firstly, set the working directory, if necessary:
## setwd("C:/Users/**********your directory***********/RepData/RepData_PeerAssessment1")

## We load the activity-data:
#############################
## 3 columns: steps date interval
ACT <- read.csv("activity.csv", header=T, colClasses=c("integer", "Date", "integer"))
Sys.setlocale(category = "LC_TIME", locale = "US")
```

## What is mean total number of steps taken per day?
(1) Firstly, we do the histogram:
```{r echo=TRUE}
stepsPerDay <- unlist(tapply(ACT[,1], ACT[,2], sum))
hist(stepsPerDay, xlab="Steps per Day", main="Total Number of Steps per Day")
abline(v=mean(stepsPerDay, na.rm=T), col="red")
abline(v=median(stepsPerDay, na.rm=T), col="blue")
```
(2) So the mean (red line) is
```{r echo=TRUE}
mean(stepsPerDay, na.rm=T)
```
and the median (blue line) is
```{r echo=TRUE}
median(stepsPerDay, na.rm=T)
```

## What is the average daily activity pattern?
(1) We do a time series plot:
```{r echo=TRUE}
averagePerIntervall <- unlist(tapply(ACT[,1], ACT[,3], mean, na.rm=T))
plot(names(averagePerIntervall), averagePerIntervall, type="l",
     xlab="Intervall Number", lwd=2, col="blue")
```

(2) The maximum number of steps across all days is in the following 5-minute intervall (begin of interval coded as "hour minute" hhmm):
```{r echo=TRUE}
maxi <- max(averagePerIntervall)
logo <- averagePerIntervall==maxi
names(averagePerIntervall[logo])
```

The maximum number of steps is (perhaps it may be interesting)
```{r echo=TRUE}
maxi
```

## Imputing missing values

(1) The number of rows with missing values is:
```{r echo=TRUE}
nrow(ACT) - sum(complete.cases(ACT))
```
(2) One possible strategy to get rid of the NA's is to substitute them with an adequate mean-value. Such a value is the mean of steps per interval.

(3) Now we substitute the missing values with this mean of steps of the intervals. Additionally we create a new dataset ACTmerged:
```{r echo=TRUE}
abc <- data.frame(interval=as.integer(names(averagePerIntervall)), steps1=averagePerIntervall)
ACTmerged <- merge(ACT, abc, by.x="interval", by.y="interval", all=T)
ACTmerged$steps <- as.numeric(ACTmerged$steps)
ACTmerged$steps[is.na(ACTmerged$steps)] <- ACTmerged$steps1[is.na(ACT[,1])]
```

(4) Now we have the following histogram:
```{r echo=TRUE}
stepsPerDay1 <- unlist(tapply(ACTmerged$steps, ACTmerged$date, sum))
hist(stepsPerDay1, xlab="Steps per Day", main="Total Number of Steps per Day")
abline(v=mean(stepsPerDay1), col="red")
abline(v=median(stepsPerDay1), col="blue")
```

So the mean (red line) is now
```{r echo=TRUE}
mean(stepsPerDay1)
```
and the median (blue line) is
```{r echo=TRUE}
median(stepsPerDay1)
```
The values are greater than before, but they are also more reasonable, because the NA's are now represented with a mean.

## Are there differences in activity patterns between weekdays and weekends?

(1) We create a factor variable with 2 levels, weekday and weekend:
```{r echo=TRUE}
dayCategoryBool <- weekdays(ACTmerged$date)=="Saturday" | weekdays(ACTmerged$date)=="Sunday"
dayCategory <- seq_along(dayCategoryBool)
dayCategory[dayCategoryBool] <- "weekend"
dayCategory[!dayCategoryBool] <- "weekday"
ACTmerged$dayCat <- as.factor(dayCategory)
```
(2) Now we do the panel plot:
```{r echo=TRUE}
weekdays1 <- ACTmerged$dayCat == "weekday"
weekend1  <- ACTmerged$dayCat == "weekend"
ACTweekdays <- ACTmerged[weekdays1,]
ACTweekend  <- ACTmerged[weekend1,]

dataweekdays <- unlist(tapply(ACTweekdays$steps, ACTweekdays$interval, mean))
dataweekend  <- unlist(tapply(ACTweekend$steps, ACTweekend$interval, mean))

df1 <- data.frame(interval=names(dataweekdays), steps=dataweekdays, mode="weekday")
df2 <- data.frame(interval=names(dataweekend), steps=dataweekend, mode="weekend")
df <- rbind(df1, df2)
df$interval <- as.numeric(as.character(df$interval))

library(lattice)
xyplot(steps ~ interval | mode, df, type="l",
       xlab="Interval", ylab="Number of steps", layout=c(1,2))
```

As can be seen, there are some differences between weekdays and weekend-days.



